### 2013-06-05 doc/pgas13/evaluation.tex
Add descriptions of diagrams, plots.

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/stack_perf.pdf}
  \caption{Stack performance.}
  \label{fig:stack}
  \end{subfigure}%
  \hspace{0.05\textwidth}
  % ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
  %(or a blank line to force the subfigure onto a new line)
  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/queue_perf.pdf}
  \caption{Queue performance.}
  \label{fig:queue}
  \end{subfigure}

  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/hashset_perf.pdf}
  \caption{GlobalHashSet performance.}
  \label{fig:hashset}
  \end{subfigure}
  %
  \hspace{0.05\textwidth}
  %
  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/hashmap_perf.pdf}
  \caption{GlobalHashMap performance.}
  \label{fig:hashmap}
  \end{subfigure}%
  %
  \caption{Raw performance of global data structures.
    \TODO{describe results.}
    \TODO{either all with error bars or none.}
  }\label{fig:datastructs}
\end{figure*}

\paragraph{Queue and Stack}
The GlobalQueue and GlobalStack have very similar implementations in terms of how they are synchronized. Both benefit greatly from flat combining. This is because only one round-trip message

However, due to the nature of concurrent stacks, the stack is able to match up pushes and pops locally. This leads to performance on an even mix of pushes and pops far outstripping the case where all pushes are performed due to requiring vastly less communication.

\paragraph{HashSet and HashMap}
The GlobalHashSet and GlobalHashMap have the same synchronization strategy (serialization happens at each hashed location).

@@

\paragraph{Breadth-First Search}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/bfs_perf.pdf}
  \caption{BFS on a Graph500-spec graph of scale 26 (64 million vertices, 1 billion edges), with the direction-optimizing BFS algorithm. Performance is shown in millions of Traversed Edges Per Second (MTEPS).}
  \label{fig:bfs_perf}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/cc_perf.pdf}
  \caption{Connected Components on the same scale 26 Graph500 graph. Performance is measured in TEPS, parameterized only by the total number of edges in the graph, regardless of the algorithm used.}
  \label{fig:cc_perf}
\end{figure}

>>>>

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/stack_perf.pdf}
  \caption{\emph{GlobalStack.} Even with Grappa's aggregation, the stack without combining completely fails to scale. With combining, both workloads scale out to 64 nodes. The workload with roughly even pushes and pops performs best because it can do matching locally.}
  \label{fig:stack}
  \end{subfigure}%
  \hspace{0.05\textwidth}
  % ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
  %(or a blank line to force the subfigure onto a new line)
  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/queue_perf.pdf}
  \caption{\emph{GlobalQueue.} Same parameters as stack performance results. The queue is unable to do matching locally, but benefits from reducing the amount of synchronization that must globally serialize. The mixed workload performs worse because the current implementation serializes combined enqueue and dequeue operations.}
  \label{fig:queue}
  \end{subfigure}

  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/hashset_perf.pdf}
  \caption{\emph{GlobalHashSet.} Throughput workload inserting and looking up random keys in the range 0-1024 into a global hash set with 1024 cells.
   Performance without combining scales better than the global queue or stack because synchronization happens at each hash cell, but drops off as number of destinations increases. Combining version is able to match inserts and lookups of the same key.}
    \TODO{if time: add 'delete' operation, too (show that it doesn't affect correctness or performance)}
  \label{fig:hashset}
  \end{subfigure}
  %
  \hspace{0.05\textwidth}
  %
  \begin{subfigure}[b]{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{data/plots/hashmap_perf.pdf}
  \caption{\emph{GlobalHashMap.} Same workload as for the Set, but random integers as the values in the map. Performance matches that of the Set.} \TODO{try increasing size of value "payload", currently is just a tiny int64.}
  \label{fig:hashmap}
  \end{subfigure}%
  %
  \caption{
    \emph{Raw performance of global data structures on simple throughput workloads.}
    Results shown for workloads with all writes and a mix of writing and reading operations. All results are for 2048 workers per core and 16 cores per node.
    \TODO{either all with error bars or none?}
    \TODO{combine stack/queue \& set/map plots, no point in saying the same things twice.}
  }\label{fig:datastructs}
\end{figure*}

\paragraph{Queue and Stack}
The GlobalQueue and GlobalStack have very similar implementations in terms of how they are synchronized. Both benefit greatly from flat combining. This is because only one round-trip message \TODO{...}

% However, the stack is able to match up pushes and pops locally. This leads to performance on an even mix of pushes and pops far outstripping the case where all pushes are performed due to requiring vastly less communication.

\paragraph{HashSet and HashMap}
The GlobalHashSet and GlobalHashMap have the same synchronization strategy (serialization happens at each hashed location).
\TODO{...}

@@

\paragraph{Breadth-First Search}
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/bfs_perf.pdf}
  \caption{\emph{BFS} on a Graph500-spec graph of scale 26 (64 million vertices, 1 billion edges), with the direction-optimizing BFS algorithm. Performance is measured in millions of Traversed Edges Per Second (MTEPS).}
  \label{fig:bfs_perf}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{data/plots/cc_perf.pdf}
  \caption{\emph{Connected Components} on the same scale 26 Graph500 graph. Performance is measured in MTEPS (parameterized by the total number of edges in the graph, independent of the algorithm used).}
  \label{fig:cc_perf}
\end{figure}

### 2013-06-05 doc/pgas13/flat-combining.tex
Add diagram of stack w/ FC, update one w/o
Add descriptions of diagrams, plots.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/fc_shared_mem.pdf}
  \caption{Flat combining in shared memory. \TODO{walkthrough...}}
  \label{fig:fc_shared_mem}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/stack.pdf}
  \caption{\emph{GlobalStack in Grappa.} Each access must synchronize with the head node (blocking if another worker is performing an operation), put or get its value in the stack's storage, release the lock on the master, and finally continue.}
  \label{fig:stack}
\end{figure}

>>>>

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/fc_shared_mem.pdf}
  \caption{\emph{Flat combining in shared memory.}
    To access the shared stack, each thread adds its request to the publication list. Thread 4 acquires the stack's lock and becomes the combiner, while the remaining threads spin waiting for their request to be satisfied. The combiner walks the publication list from the head, matching up Thread 3's push with Thread 2's pop on its own private stack. The 2 remaining pushes are added to the shared stack, the top is updated, and finally Thread 4 releases the lock and continues its normal execution.
  }
  \label{fig:fc_shared_mem}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.43\textwidth}
    \includegraphics[width=\textwidth]{figs/stack_nofc.pdf}
    \caption{\emph{Without flat combining.}
      To do its push, Worker 1 sends a message to synchronize with the master on Core 0, sends another message to write its value to the top of the stack, bumps the synchronized \texttt{top} pointer, and finally continues. Worker 2, and all other workers on all cores accessing the stack, must block at the master and wait for Worker 1 to complete its push before doing their operations.
    }
    \label{fig:stacknofc}
  \end{subfigure}%
  \hspace{0.05\textwidth}
  \begin{subfigure}[b]{0.43\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/stack_fc.pdf}
    \caption{\emph{Local flat combining.}
      Workers within the core add their requests to a local proxy object. Worker 3's pop matches with Worker 2's push, requiring no global communication. After combining locally, worker 1 and 4's pushes remain, so Worker 4 becomes the core's combiner, synchronizes with the master, adds both new values to the global stack, bumps the top pointer and releases the lock on master, and finally wakes Worker 1 and continues.
    }
    \label{fig:stackfc}
  \end{subfigure}
  \caption{\emph{Global stack in Grappa} with and without \emph{flat combining}.}
  \label{fig:stack}
\end{figure*}

### 2013-06-05 doc/pgas13/grappa.tex
Work on flat-combining section (also fix \TODO, add fc papers to bib)
First draft of grappa,fc_shared_mem, and non_fc_stack
Add descriptions of diagrams, plots.

\section{Grappa}
Existing shared memory applications typically do fine-grained operations on data, assuming it is all cheap to access.
Naively translating these applications to PGAS languages often results in poor performance because many operations must then be done via fine-grained remote communication.
Much research has been done to improve performance of PGAS implementations by coming up with better ways to partition data and schedule computation, and do communication in bulk to best utilize network resources.
For regular workloads, most pointedly dense linear algebra, much of this can now be automatically optimized generically.

@@

\subsection{Tasks}
\todo{describe how tasks work}

\subsection{Global Memory}
\todo{explain how memory is partitioned}

\subsection{Communication}
\todo{remote delegate operations, aggregation}

>>>>

\section{Grappa}
Existing shared memory applications typically do fine-grained operations on data, assuming it is all cheap to access.
Naively translating these applications to PGAS languages often results in poor performance because many operations must then be done via fine-grained remote communication.
Much research has been done to improve performance of PGAS implementations by coming up with better ways to partition data and schedule computation, and do communication in bulk to best utilize network resources.
For workloads with regular access patterns, for instance dense linear algebra or structured grids, there are proven techniques that can automatically optimize communication effectively. \TODO{cite or remove prev. statement}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/grappa_system.pdf}
  \caption{\emph{Grappa System Overview:}
    Thousands of workers are multiplexed onto each core, context switching to tolerate the additional latency of aggregating remote operations. Each core has exclusive access to a slice of the global heap which is partitioned across all nodes, as well as core-private data and worker stacks. Tasks can be spawned and executed anywhere, but are bound to a worker in order to be executed.
  }
  \label{fig:system}
\end{figure}

\subsection{Tasks}
\TODO{describe how tasks work}

\subsection{Global Memory}
\TODO{explain how memory is partitioned}

\subsection{Communication}
\TODO{remote delegate operations, aggregation}

### 2013-06-05 doc/pgas13/intro.tex
work on flat-combining section, add todos, fix typo in intro
Add descriptions of diagrams, plots.

The trick is coming up with a mechanism for delegating that has lower synchronization overhead than the original contention case.
Prior work showed that a cleverly-implemented ``publication lists'' can allow multiple accessors to cooperate with minimal synchronization.
This same work makes the additional observation that, given the semantics of some particular data structures, the set of accesses being serialized can be composed and performed more efficiently \emph{combined} than individually, which they dub \emph{flat combining}.
Together, the reduced synchronization cost and combined operations allow even a data structure with a single global lock to scale better than complicated concurrent data structure implementations with fine-grained locking.
Further, the same publication list mechanism can be applied to other data structures, and all that must be customized is the particular way in which operations for a given data structure can be combined.

@@

The goal of this work is to apply the concept of flat-combining in a PGAS runtime to reduce the cost of maintaining globally consistent data structures in a generic way.
To evaluate this, we leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
With enough concurrent threads, it is actually possible to maintain program order within each thread by blocking on accesses, but still have an opportunity to effectively combine many fine-grained synchronous operations for better performance.
With the same flat-combining mechanism, multiple global data structures can be implemented efficiently, even with extremely simple locking schemes.

% lay out the upcoming sections of the paper...
The next section will describe in more detail the Grappa runtime system that is used to implement flat combining for distributed memory machines. Then we explain the flat combining mechanism in more depth and describe how it maps to a PGAS model. Next we explain how several data structures are implemented in this framework and show how they perform on simple throughput workloads. Finally, we evaluate how they affect performance when integrated into a couple application kernels.

>>>>

The trick is coming up with a mechanism for delegating that has lower synchronization overhead than the original contention case.
Prior work showed that a cleverly-implemented ``publication list'' can allow multiple accessors to cooperate with minimal synchronization.
This same work makes the additional observation that, given the semantics of some particular data structures, the set of accesses being serialized can be composed and performed more efficiently \emph{combined} than individually, which they dub \emph{flat combining}.
Together, the reduced synchronization cost and combined operations allow even a data structure with a single global lock to scale better than complicated concurrent data structure implementations with fine-grained locking.
Further, the same publication list mechanism can be applied to other data structures, and all that must be customized is the particular way in which operations for a given data structure can be combined.

@@

The goal of this work is to apply the concept of flat-combining in a PGAS runtime to reduce the cost of maintaining globally consistent data structures in a generic way.
To evaluate this, we leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
With enough concurrent threads, it is actually possible to maintain program order within each thread by blocking on accesses, but still have an opportunity to effectively combine many fine-grained synchronous operations for better performance.
With the same flat-combining mechanism, multiple global data structures can be implemented efficiently, even with simple locking schemes.

% lay out the upcoming sections of the paper...
The next section will describe in more detail the Grappa runtime system that is used to implement flat combining for distributed memory machines. Then we explain the flat combining mechanism in more depth and describe how it maps to a PGAS model. Next we explain how several data structures are implemented in this framework and show how they perform on simple throughput workloads. Finally, we evaluate how they affect performance when used in a simple application kernels.
